2-D Photosynthesis Model

### 2023-09-26 ms/si.qmd
corrected f_liq calculation following Tom's email on 2023-09-26

As I worked through @earles_excess_2017 model, it did not seem like dividing $g_\text{liq}$ by $l_z$ in equation @eq-fliq1 made sense. The FEM is a discretization of a continuous model. Intuitively, it does make sense why a coarser grid (greater $l_z$) would lead to a larger \[CO$_2$\] drawdown for a given assimilation rate. Noting that $g_\text{liq}$ is conductance per m$^2$ of stroma, I think the correct way is to divide by $V_\text{strom}$, which gives the mesophyll surface area per stroma volume available for liquid-phase diffusion. Here I assume that the chloroplasts line the entire inner wall of the mesophyll, so 1 [m$^2$] mesophyll = 1 [m$^2$] stroma, but this could be modified to the correct value. My equation is therefore:

$$f_\text{liq} = \frac{g_\text{liq} (C_\text{liq} - C_\text{ias})}{V_\text{strom}}$$ {#eq-fliq2}

Here's how I work out the units for equation @eq-fliq2.

$$\frac{\text{mol}}{\text{m}^3~\text{s}} = \bigg(\frac{\text{m}^3}{\text{m}^2~\text{stroma}} \times \frac{1~\text{m}^2~\text{stroma}}{1~\text{m}^2~\text{meso}} \bigg) \times \bigg(\frac{\text{mol}}{\text{m}^3} \bigg) \times \bigg( \frac{\text{m}^2~\text{meso}}{\text{m}^3} \bigg)$$

This still doesn't seem quite right to me because it seems like the change in airspace concentration should be affected by the porosity since a smaller volume of air will experience a greater drop in concentration for the same amount of assimilation, but maybe I am thinking about this incorrectly.

Coincidentally, this had little effect on @earles_excess_2017 model because their element size was $\approx 1~\mu$m, which is quantitatively similar to $V_\text{strom} = 1.74 \times 10^{-6}$m. However, when you change the element size on their model, the solution is quite different; when you change the element size on my model, the solution is only slightly affected but the discretization is an approximation of the true solution.

@@

@@

  # diffusion in Z-direction
  Flux = -D_e / parms[["t_elem"]] * rbind(
    C_ias[1, ] - bound_bottom,
    (C_ias[2:parms[["n_z"]], ] - C_ias[1:(parms[["n_z"]] - 1), ]),
    bound_top - C_ias[parms[["n_z"]], ]
  )
  dC_ias = dC_ias -
    (Flux[2:(parms[["n_z"]] + 1), ] - Flux[1:parms[["n_z"]], ]) / parms[["t_elem"]]

  dC_ias = dC_ias + parms[["g_liq"]] * (C_liq - C_ias) / parms[["V_strom"]]

  # CARBOXYLATION ----
  # Leaf thickness - needed to scale from stroma volume per leaf volume
  T_leaf = parms[["n_z"]] * parms[["t_elem"]] # [m]
  dC_liq = dC_liq + parms[["g_liq"]] * (C_ias - C_liq) / parms[["V_strom"]] +
    (-r_c + r_p + r_d) * (parms[["S_m"]] * 1 / T_leaf) * parms[["V_strom"]]

>>>>

As I worked through @earles_excess_2017 model, it did not seem like dividing $g_\text{liq}$ by $l_z$ in equation @eq-fliq1 made sense. The FEM is a discretization of a continuous model. Intuitively, it does make sense why a coarser grid (greater $l_z$) would lead to a larger \[CO$_2$\] drawdown for a given assimilation rate. Noting that $g_\text{liq}$ is conductance per m$^2$ of stroma, this means the length scale to divide by should be 1/(stroma area per unit bulk leaf volume). This is equivalent to 1/\[(stroma area per leaf area)x(leaf area per bulk leaf volume)\] = 1/\[Sc x (1/Tleaf)\] = Tleaf/Sc (for now I am assuming Sm = Sc, but we can correct this later). Our equation is therefore:

$$f_\text{liq} = \frac{g_\text{liq} (C_\text{liq} - C_\text{ias})}{T_\text{leaf} / S_\text{m}}$$ {#eq-fliq2}

@@

  # diffusion in Z-direction; boundaries=imposed concentration
  Flux = -D_e / parms[["t_elem"]] * rbind(
    C_ias[1, ] - bound_bottom,
    (C_ias[2:parms[["n_z"]], ] - C_ias[1:(parms[["n_z"]] - 1), ]),
    bound_top - C_ias[parms[["n_z"]], ]
  )
  dC_ias = dC_ias -
    (Flux[2:(parms[["n_z"]] + 1), ] - Flux[1:parms[["n_z"]], ]) / parms[["t_elem"]]

  dC_ias = dC_ias + parms[["g_liq"]] * (C_liq - C_ias) /
    (parms[["T_leaf"]] / parms[["S_m"]])

@@

  dC_liq = dC_liq +
    parms[["g_liq"]] * (C_ias - C_liq) / (parms[["T_leaf"]] / parms[["S_m"]]) +
    (-r_c + r_p + r_d) * (parms[["S_m"]] / parms[["T_leaf"]]) * parms[["V_strom"]]
